---
title: "p8105_hw6_zl3386"
author: "Ziqiu Liu"
date: "2023-11-21"
output: github_document
---

```{r, message = FALSE}
library(tidyverse)
library(rvest)
library(readr)
```

## Problem 1

```{r import data}
homicide_df = 
  read_csv("data/homicide-data.csv") |>
  mutate(
    victim_age = as.numeric(victim_age),
    state = toupper(state)
  ) |>
  mutate(
    # create a `city_state` variable
    city_state = paste(city, ",", state),
    # a binary variable indicating whether the homicide is solved
    solved_status = case_match( 
      disposition,
      "Closed by arrest" ~ 1,
      c("Closed without arrest", "Open/No arrest") ~ 0,
    )
  ) |>
  # Omit cities that don’t report victim race and Tulsa, AL
  filter(!(city_state %in% c("Dallas , TX", "Phoenix , AZ", "Kansas City , MO"))) |> 
  filter(city_state != "Tulsa , AL") |>
  # limit to white or black victim_race
  filter(victim_race %in% c("White", "Black"))
```

For the city of Baltimore, MD, use the `glm` function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors.

Save the output of `glm` as an R object; apply the `broom::tidy` to this object; and obtain the estimate and confidence interval of the adjusted **odds ratio** for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
baltimore_df = homicide_df |>
  filter(city_state == "Baltimore , MD") |>
  select(solved_status, victim_age, victim_sex, victim_race) 
```

